Results for working group papers

### 2016-06-14 evos-ifq/catch_portfolios_EVOSIFQ.Rmd
Start rearranging the files

---
title: "Results for working group papers"
author: "Catch portfolios subgroup: NCEAS meeting 3"
date: "March 18, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
```

## Load cleaned data  
```{r dataLoad, warning = FALSE, message=FALSE}
library(dplyr)
library(knitr)
library(mgcv)
library(ggplot2)
library(randomForest)
library(MASS)
library(zoo)
library(MatchIt)
load("cfecCleaned.Rdata")
```

```{r, warning = FALSE, message=FALSE}
# This is to investigate Bristol Bay set nets for Milo
# Check BB setnet records / year
#idx = grep("S 04T", cfec$p_fshy)
#group_by(cfec[idx,], year) %>% 
#  summarize(nSetnet = length(unique(p_holder)))
#cfecnew = cfec[idx,]
```

## Function definiton  
We need to define a few functions for calculating diversity. We'll use the Simpson's diversity index (HHH) to calculate the entropy-based inverse, which is a proxy for effective species diversity. Note the C++ version is commented out below,   
```{r, warning = FALSE, message=FALSE}

simp.div = function(x) {
  1/sum((x/sum(x))^2)
}

#library(Rcpp)
#cppFunction('double simpCpp(NumericVector x) {
#  double y = 1/sum((x/sum(x))^2);
#  return y;
#}')
```
# Paper 1:

To start, we can coarsely summarize the variability in total revenue by species through time, 
```{r, fig.pos="placeHere", fig.cap = "Log(total) revenue of each spp"}
spTot = group_by(cfec, specn, year) %>%
  summarize(totRev.bySpec = log(sum(g_earn, na.rm=T))) %>% 
  as.data.frame() %>%
  filter(is.finite(totRev.bySpec)) %>%   
  as.data.frame
ggplot(spTot, aes(x=year,y=totRev.bySpec,group=as.factor(specn))) +geom_line(aes(x=year,y=totRev.bySpec, color=as.factor(specn)))
```

```{r, fig.pos="placeHere", fig.cap = "CV of log(total revenue) across species (sd in log space)"}
spTot = group_by(spTot, year) %>%
  summarize(sd_yr = sd(totRev.bySpec, na.rm=T)) %>% 
  as.data.frame()
ggplot(spTot, aes(x=year,y=sd_yr)) + geom_point()

@@

For each individual in each year, we can calculate the effective number of vessels used, and trends in that (mean across people).   
```{r}
# calculate diversity of individuals across vessels / time
cfec.boatDiv = group_by(cfec, p_holder, year) %>% 
  summarize(effBoats = 1 / sum((table(cadfg) / sum(table(cadfg)))^2) ) %>% 
  data.frame

cfec.boatDiv = as.data.frame(cfec.boatDiv)
g = gam(as.numeric(effBoats) ~ s(as.numeric(year)), data=cfec.boatDiv)
gpred = predict(g, newdata=list(year=1985:2014), se.fit=TRUE)
plot(1985:2014, gpred$fit, type="l", lwd=3, ylim=c(1.03,1.1))
lines(1985:2014, gpred$fit+2*gpred$se.fit)
lines(1985:2014, gpred$fit-2*gpred$se.fit)
```

Now we'll filter out individuals who only used 1 boat in a year, and boats with only one permit holder fishing in a year.  
```{r}
cfecnew = cfec
cfec.1boat = group_by(cfecnew, p_holder, year) %>% 
  summarize(nBoat = length(unique(cadfg))) %>%
  filter(nBoat==1) %>% 
  left_join(cfecnew)

@@

So there's about 50% of the records remaining -- these are unique individial / vessel combinations. Side note: all of the set net / razor clam permits and individuals have been included. Not sure we want to drop them out [Milo - thoughts? ]

Now we can create a unique *id* column for individual - vessel combinations.

```{r}
cfecnew$id = paste(cfecnew$p_holder,cfecnew$cadfg)
# calculate number of unique person - vessel pairs ~ 40700
length(unique(cfecnew$id))
```

Just for demo purposes, we can start by plotting mean - CV relationships for each of these individual- vessel combinations (across years).

```{r, fig.pos="placeHere", fig.cap="Mean and variability (CV) of annual earnings (across time) for unique people-vessel combinations."}
meanVar = group_by(cfecnew, id, year) %>% 
  summarize(totRev = sum(g_earn)) %>% 
  as.data.frame() %>%
  group_by(id) %>%
  summarize(m = mean(log(totRev+1),na.rm=T), cv = sd(totRev,na.rm=T)/mean(totRev,na.rm=T))
plot(meanVar$m, meanVar$cv, xlab="Mean log(annual revenue)", ylab = "CV annual revenue")
```

One question we also were chatting about was what the distribution of the percent change metric looked like. 
```{r mergeYears}

totalcfec = group_by(cfecnew, id, year) %>% 
  summarize(totrev = sum(g_earn))

totalcfec$year = as.numeric(totalcfec$year)
dummyDF = expand.grid(id = unique(totalcfec$id), 
  year = min(totalcfec$year) : max(totalcfec$year), 
  stringsAsFactors = F)
# expand the data frame w/missing years
cfecExp = left_join(dummyDF, totalcfec) %>%
  arrange(id, year)
cfecExp = group_by(cfecExp, id) %>%
  mutate(difflog = c(NA, diff(log(totrev))), meanRev = c(NA, zoo::rollmean(totrev, 2))) %>%
  filter(!is.na(difflog))
```

This shows the pattern we'd hope to expect -- no trend as a function of mean revenue, . Namely that people making less than about 35K / year tend to have negative returns from one year to the next, and people with revenues > 35K -- especially the high earners -- tend to have positive returns. We also see decreased variability as a function of returns -- maybe because investments in technology, permits, etc.
```{r, fig.pos="placeHere", fig.cap = "Distribution of percent change in total revenue + relationship (or lack of) with total annual revenue"}
par(mfrow = c(2,1), mai=c(0.5,0.5,0.1,0.1))
plot(log(cfecExp$meanRev), cfecExp$difflog, xlab = "log(revenue)", ylab="percent change in revenue", cex=0.7)
abline(0, 0, col="red")

hist(cfecExp$difflog, 80, xlab="Percent return (empirical)", col="grey", xlim= quantile(cfecExp$difflog, c(.005,0.995)), main="")

@@

```{r, warning = FALSE, message=FALSE}
# We'll use 3 years before/after IFQ. This is consistent with the work that
# Alan has done. Years up to 1990 were included in the calculation of future
# quota, and 1998 is problematic because the ratio of our halibut earnings to 
# AKFIN's is ~ 60% -- suggesting there may be something systemic that we're
# missing. 
cfec$year = as.numeric(cfec$year)

@@

```{r, cache = TRUE, warning = FALSE, message=FALSE}

@@

```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by fish ticket freq.) through time for each of our 4 groups during the pre-post IFQ period. "}
ggplot(hs.day.summary, aes(year, m.freq)) + geom_line() + facet_wrap(~group) + 
  geom_ribbon(aes(ymin=l.freq, ymax=u.freq), alpha=0.2)
```

```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by earnings) through time for each of our 4 groups during the pre-post IFQ period. "}
ggplot(hs.day.summary, aes(year, m.earn)) + geom_line() + facet_wrap(~group) + 
  geom_ribbon(aes(ymin=l.earn, ymax=u.earn), alpha=0.2)
```

```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by pounds of catch) through time for each of our 4 groups during the pre-post IFQ period. "}
ggplot(hs.day.summary, aes(year, m.weight)) + geom_line() + facet_wrap(~group) + 
  geom_ribbon(aes(ymin=l.weight, ymax=u.weight), alpha=0.2)
```

@@

First, make a dotplot (with quartile range) of the pre-post comparison for each metric of revenue.  
```{r, fig.pos="placeHere", fig.cap = "Metric of annual revenue, with interquartile range, for each of the 4 IFQ groups."}

@@

Look at the mean number of days where catch was recorded as landed, as well as how many of those days are when halibut are landed.  
```{r}
dplyr::select(hs.summary, group, ind, p_holder, 
  contains("nDays"), contains("nHalibutDays")) %>%
  reshape2::melt(id.vars = c("group", "ind","p_holder")) %>%
  group_by(group, ind, variable) %>% 
  filter(is.na(value)==F) %>% 
  summarize(l = quantile(value, 0.25, na.rm=T), 
    u = quantile(value, 0.75, na.rm=T), 
    m = quantile(value, 0.5, na.rm=T)) %>%
  ggplot(aes(x = ind, y = m, ymax = u, ymin = l)) + 
  geom_pointrange() + facet_grid(variable ~ group, scales="free")
```

```{r, fig.pos="placeHere", fig.cap = "Boxplots of revenue metrics (with interquartile range) by group and pre-post-IFQ. ", include=FALSE, echo=FALSE, message = FALSE, warning = FALSE}
#dplyr::select(hs.summary, group, ind, p_holder, contains("mean"), -contains("div")) %>% 
#  reshape2::melt(id.vars = c("group", "ind","p_holder")) %>%
#  ggplot(aes(ind, value)) + geom_boxplot(aes(fill=ind), alpha=0.5, outlier.size=NA) + 
#  facet_grid(variable ~ group, scales="free")
```

```{r, fig.pos="placeHere", fig.cap = "Boxplots of diversity metrics (with interquartile range) by group and pre-post-IFQ. "}
dplyr::select(hs.summary, group, ind, p_holder, contains("eff.spec.div.mean"), contains("eff.gear.div.mean"), contains("eff.spec.div.mean")) %>% 
  reshape2::melt(id.vars = c("group", "ind","p_holder")) %>%
  ggplot(aes(ind, value)) + 
  geom_boxplot(aes(fill=ind), alpha=0.5, outlier.size=NA) + 
  facet_grid(variable ~ group, scales="free")
# +  geom_point(alpha = 0.02, position = position_jitter(width = 0.4), cex = 0.2)

@@

```{r, fig.pos="placeHere", fig.cap="Violin plot of individual differences (post-IFQ - pre-IFQ for the revenue and diversity metrics included in our analysis."}
hs.long = reshape2::melt(hs.summary, id.vars = c("group", "ind","p_holder"))
hs.long = reshape2::dcast(hs.long, variable + p_holder + group ~ ind) %>%
  mutate(diff_ = (Post-Pre)/Pre)

@@

```{r, fig.pos = "placeHere", fig.cap = "Gear diversity through time for each of the 3 groups, calculated based on the catch (black) and earnings (red)."}
hs.summary.gear = group_by(dat, ind, group, year, gearn) %>% 
  summarize(n = n(), nEarn = sum(g_earn), nWeight = sum(g_pounds)) %>% 
  as.data.frame %>% 
  group_by(ind, group, year) %>% 
  summarize(gearDiv.freq = simp.div(n), 
    gearDiv.earn = simp.div(nEarn), 
    gearDiv.weight = simp.div(nWeight))

@@

We'll break down the proportion used for the top 8 gears,
```{r, fig.pos="placeHere", fig.cap = "Time series of proportion of gears used"}
hs.gear.tot = group_by(dat, ind, group, year) %>% 
  summarize(nTot = n())
hs.summary.gear = group_by(dat, ind, group, year, gearn) %>% 
  summarize(n = n()) %>%
  left_join(hs.gear.tot)
hs.summary.gear$p = hs.summary.gear$n/hs.summary.gear$nTot
hs.summary.gear = filter(hs.summary.gear, gearn %in% c(1,6,15,7,9,47,4,5)) %>% 
ggplot(aes(year, p)) + geom_line() + facet_grid(gearn ~ group, scales="free")

@@

Can we predict effective species diversity as function of gear diversity?
```{r}
#hs.geardiv = group_by(dat, ind, group, p_holder, year, gearn) %>% 
#  summarize(tot.weight.sp = sum(g_pounds), 
#    tot.revenue.sp = sum(g_earn)) %>% 
#  group_by(ind, group, p_holder, year) %>% 
#  summarize(eff.gear.div.mean.earn = simp.div(tot.revenue.sp),
#    eff.gear.div.mean.weight = simp.div(tot.weight.sp)) %>%
#  group_by(ind, group, p_holder) %>% 
#  summarize(eff.gear.div.cv.earn = 
#      sd(eff.gear.div.mean.earn) / mean(eff.gear.div.mean.earn), 
#    eff.gear.div.mean.earn = mean(eff.gear.div.mean.earn), 
#    eff.gear.div.cv.weight = 
#      sd(eff.gear.div.mean.weight) / mean(eff.gear.div.mean.weight),
#    eff.gear.div.mean.weight = mean(eff.gear.div.mean.weight))
#hs.summary = left_join(hs.summary, hs.geardiv) %>% 
#  as.data.frame()
```

@@

We'll restrict our analysis initially to a few years before and after EVOS (1989). To make the comparison similar to the IFQ comparison, we will define the following groups:

1. Individuals who fished in closed areas before and after spill  
2. Individuals who fished in closed areas before spill ["Pre"]
3. Individuals who only fished in closed areas after 1995 ["Post"]  
4. Individuals who never fished in closed areas ["Control group"]

## Propensity matching for EVOS
We're interested in forming a control group for the comparison of the EVOS effects...
```{r, warning = FALSE, message=FALSE}

# Don't include 1989, but include before / after
dat = cfec[cfec$year %in% c(1985:1988, 1990:1991),]
dat$day = substr(dat$landdate, 6, 10)
# Create an indicator based on the salmon / shellfish araes that were closed in PWS
dat$exposed = 0
dat$exposed[dat$stat6%in%c(22710, 22720, 22730, 22610, 22620, 22630, 22640, 22650, 22510, 22520, 22530)] = 1

# Exposed variable is the response, and we want to predict based on traits
dat.ev = group_by(dat, p_holder, year) %>% 
  summarize(tot.revenue.ind = sum(g_earn), 
    tot.revenue.ind.her = sum(g_earn[specn==230]), 
    tot.revenue.ind.sal = sum(g_earn[specn%in%c(400,410,420,430,440,450)]),
    tot.pounds.ind = sum(g_pounds), 
    tot.pounds.ind.her = sum(g_pounds[specn==230]), 
    tot.pounds.ind.sal = sum(g_pounds[specn%in%c(400,410,420,430,440,450)]),
    eff.spec.div = simp.div(table(specn)), 
    eff.gear.div = simp.div(table(gearn)),
    numDays = length(unique(day)),
    exposed = max(exposed), 
    ind = ifelse(year[1] < 1989, "Pre", "Post")) %>% 
    as.data.frame() %>%
  group_by(p_holder, ind) %>% 
  summarize(tot.revenue.ind = mean(tot.revenue.ind), 
    tot.revenue.ind.her = mean(tot.revenue.ind.her), 
    tot.revenue.ind.sal = mean(tot.revenue.ind.sal),
    tot.pounds.ind = mean(tot.pounds.ind), 
    tot.pounds.ind.her = mean(tot.pounds.ind.her), 
    tot.pounds.ind.sal = mean(tot.pounds.ind.sal),
    eff.spec.div = mean(eff.spec.div), 
    eff.gear.div = mean(eff.gear.div),
    exposed = max(exposed),
    meanDays = mean(numDays)) %>% 
    as.data.frame()

@@

@@

dat.ev.rf[,3:8] = log(dat.ev.rf[,3:8] + 1)

m.glm = glm(exposed ~ ., data = dat.ev.rf[dat.ev.rf$ind=="Pre", -c(1,2)], family = "binomial")
stepAIC(m.glm, direction="both")

bestModel = glm(exposed ~ tot.revenue.ind.her + tot.revenue.ind.sal + 
    tot.pounds.ind + tot.pounds.ind.her + tot.pounds.ind.sal + 
    eff.spec.div + meanDays + eff.spec.div.mean.weight, data = 
    dat.ev.rf[dat.ev.rf$ind == "Pre", ])

m.out <- matchit(exposed ~ tot.revenue.ind + tot.revenue.ind.her + 
    tot.revenue.ind.sal + tot.pounds.ind + tot.pounds.ind.her + 
    tot.pounds.ind.sal + eff.spec.div + eff.gear.div + meanDays + 
    eff.spec.div.mean.earn, method = "nearest", data = dat.ev.rf[dat.ev.rf$ind == 
    "Pre", ])

@@

```{r}

>>>>

---
title: "Results for working group papers"
author: 'Catch portfolios subgroup: NCEAS meeting 3'
date: "March 18, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document: default
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
```

## Load cleaned data  
```{r dataLoad, warning = FALSE, message=FALSE, results="hide"}
library(dplyr)
library(knitr)
library(mgcv)
library(ggplot2)
library(randomForest)
library(MASS)
library(zoo)
library(MatchIt)
load("cfecCleaned2.Rdata")
```

We need to adjust the landed values and prices for inflation. To do this, we'll use the data Alan pulled together (adjusting to 2009 dollars).

https://docs.google.com/spreadsheets/d/1hRLYTX1nZBLKnYXwTCOtU7-XnDoMZf_1ufdTiNQFd4U/edit#gid=0

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
deflationTable = data.frame(year = 1985:2014, defl = c(0.57236, 0.58393, 0.59879, 0.61974,  
0.64388, 0.66774, 0.68993, 0.70564, 0.72244, 0.73781,  
0.75321, 0.76695, 0.78009, 0.78855, 0.80061,  
0.81883, 0.83753, 0.85038, 0.86729, 0.89114,  
0.91981, 0.94812, 0.97334, 0.9925, 1,  
1.01217, 1.03307, 1.05213, 1.06926, 1.08682))

# Adjust price and g_earn for inflation
cfec$year = as.numeric(cfec$year)
cfec$g_price = cfec$g_price/deflationTable$defl[cfec$year-1984]
cfec$g_earn = cfec$g_earn/deflationTable$defl[cfec$year-1984]

simp.div = function(x) {
  1/sum((x/sum(x))^2)
}
```

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
cfecnew = cfec
cfec.1boat = group_by(cfecnew, p_holder, year) %>% 
  summarize(nBoat = length(unique(cadfg))) %>%
  filter(nBoat==1) %>% 
  left_join(cfecnew)

@@

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
# We'll use 3 years before/after IFQ. This is consistent with the work that
# Alan has done. Years up to 1990 were included in the calculation of future
# quota, and 1998 is problematic because the ratio of our halibut earnings to 
# AKFIN's is ~ 60% -- suggesting there may be something systemic that we're
# missing. 
cfec$year = as.numeric(cfec$year)

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}

@@

```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by fish ticket freq.) through time for each of our 4 groups during the pre-post IFQ period. ", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
ggplot(hs.day.summary, aes(year, m.freq)) + geom_line() + facet_wrap(~group) + 
  geom_ribbon(aes(ymin=l.freq, ymax=u.freq), alpha=0.2)
```

```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by earnings) through time for each of our 4 groups during the pre-post IFQ period. ", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
ggplot(hs.day.summary, aes(year, m.earn)) + geom_line() + facet_wrap(~group) + 
  geom_ribbon(aes(ymin=l.earn, ymax=u.earn), alpha=0.2)
```

```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by pounds of catch) through time for each of our 4 groups during the pre-post IFQ period. ", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
ggplot(hs.day.summary, aes(year, m.weight)) + geom_line() + facet_wrap(~group) + 
  geom_ribbon(aes(ymin=l.weight, ymax=u.weight), alpha=0.2)
```

@@

First, make a dotplot (with quartile range) of the pre-post comparison for each metric of revenue.  
```{r, fig.pos="placeHere", fig.cap = "Metric of annual revenue, with interquartile range, for each of the 4 IFQ groups.", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}

@@

Look at the mean number of days where catch was recorded as landed, as well as how many of those days are when halibut are landed.  
```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
dplyr::select(hs.summary, group, ind, p_holder, 
  contains("nDays"), contains("nHalibutDays")) %>%
  reshape2::melt(id.vars = c("group", "ind","p_holder")) %>%
  group_by(group, ind, variable) %>% 
  filter(is.na(value)==F) %>% 
  summarize(l = quantile(value, 0.25, na.rm=T), 
    u = quantile(value, 0.75, na.rm=T), 
    m = quantile(value, 0.5, na.rm=T)) %>%
  ggplot(aes(x = ind, y = m, ymax = u, ymin = l)) + 
  geom_pointrange() + facet_grid(variable ~ group, scales="free")
```

```{r, fig.pos="placeHere", fig.cap = "Boxplots of revenue metrics (with interquartile range) by group and pre-post-IFQ. ", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
#dplyr::select(hs.summary, group, ind, p_holder, contains("mean"), -contains("div")) %>% 
#  reshape2::melt(id.vars = c("group", "ind","p_holder")) %>%
#  ggplot(aes(ind, value)) + geom_boxplot(aes(fill=ind), alpha=0.5, outlier.size=NA) + 
#  facet_grid(variable ~ group, scales="free")
```

```{r, fig.pos="placeHere", fig.cap = "Boxplots of diversity metrics (with interquartile range) by group and pre-post-IFQ. ", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
dplyr::select(hs.summary, group, ind, p_holder, contains("eff.spec.div.mean"), contains("eff.gear.div.mean"), contains("eff.spec.div.mean")) %>% 
  reshape2::melt(id.vars = c("group", "ind","p_holder")) %>%
  ggplot(aes(ind, value)) + 
  geom_boxplot(aes(fill=ind), alpha=0.5, outlier.size=NA) + 
  facet_grid(variable ~ group, scales="free")
# +  geom_point(alpha = 0.02, position = position_jitter(width = 0.4), cex = 0.2)

@@

```{r, fig.pos="placeHere", fig.cap="Violin plot of individual differences (post-IFQ - pre-IFQ for the revenue and diversity metrics included in our analysis.", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
hs.long = reshape2::melt(hs.summary, id.vars = c("group", "ind","p_holder"))
hs.long = reshape2::dcast(hs.long, variable + p_holder + group ~ ind) %>%
  mutate(diff_ = (Post-Pre)/Pre)

@@

```{r, fig.pos = "placeHere", fig.cap = "Gear diversity through time for each of the 3 groups, calculated based on the catch (black) and earnings (red).", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
hs.summary.gear = group_by(dat, ind, group, year, gearn) %>% 
  summarize(n = n(), nEarn = sum(g_earn), nWeight = sum(g_pounds)) %>% 
  as.data.frame %>% 
  group_by(ind, group, year) %>% 
  summarize(gearDiv.freq = simp.div(n), 
    gearDiv.earn = simp.div(nEarn), 
    gearDiv.weight = simp.div(nWeight))

@@

We'll break down the proportion used for the top 8 gears,
```{r, fig.pos="placeHere", fig.cap = "Time series of proportion of gears used", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
hs.gear.tot = group_by(dat, ind, group, year) %>% 
  summarize(nTot = n())
hs.summary.gear = group_by(dat, ind, group, year, gearn) %>% 
  summarize(n = n()) %>%
  left_join(hs.gear.tot)
hs.summary.gear$p = hs.summary.gear$n/hs.summary.gear$nTot
hs.summary.gear = filter(hs.summary.gear, gearn %in% c(1,6,15,7,9,47,4,5)) %>% 
ggplot(aes(year, p)) + geom_line() + facet_grid(gearn ~ group, scales="free")

@@

Can we predict effective species diversity as function of gear diversity?
```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
#hs.geardiv = group_by(dat, ind, group, p_holder, year, gearn) %>% 
#  summarize(tot.weight.sp = sum(g_pounds), 
#    tot.revenue.sp = sum(g_earn)) %>% 
#  group_by(ind, group, p_holder, year) %>% 
#  summarize(eff.gear.div.mean.earn = simp.div(tot.revenue.sp),
#    eff.gear.div.mean.weight = simp.div(tot.weight.sp)) %>%
#  group_by(ind, group, p_holder) %>% 
#  summarize(eff.gear.div.cv.earn = 
#      sd(eff.gear.div.mean.earn) / mean(eff.gear.div.mean.earn), 
#    eff.gear.div.mean.earn = mean(eff.gear.div.mean.earn), 
#    eff.gear.div.cv.weight = 
#      sd(eff.gear.div.mean.weight) / mean(eff.gear.div.mean.weight),
#    eff.gear.div.mean.weight = mean(eff.gear.div.mean.weight))
#hs.summary = left_join(hs.summary, hs.geardiv) %>% 
#  as.data.frame()
```

\newpage

@@

We'll restrict our analysis initially to a few years before and after EVOS (1989). To make the comparison similar to the IFQ comparison, we will have an impacted group. We'll define 'impacted' as people who were landing catch in statistical areas over 1985-1988 that were later closed because of the spill (1989). These areas were then opened up for fishing in 1990. We'll restrict the analysis to only looking at immediate impacts, so we're interested in how affected / unaffected individuals fared in 1990-1991.

To create a control group, we'll use propensity matching (other clustering methods also an option). This procedure involves:  
(1) defining attributes for all individuals during 1985-1988 (total catch, total revenue, % revenue from salmon, etc).  
(2) Doing some simple model selection with a binomial GLM to reduce the number of variables included in our model (some will be redundant with others, and others will be uninformative)  
(3) Run propensity matching to identify among the unaffected individuals a best match for each impacted individual. Note that this will only identify 1 match for each case. 
(4) Using the affected individuals and new control group of 'matches', examine how the two groups fared post spill, in 1990-1991. (As an aside, Rich thought many individuals fishing salmon, etc. in these areas were going right back to the previously closed areas).

## Propensity matching for EVOS

We're interested in forming a control group for the comparison of the EVOS effects.  To do this, we adopted the following approach:  
1.  Filter the data to be only 1985:1988, and 1990:1991 as our pre/post group. Only use data for people who fished before and after.    
2.  Create an 'exposed' treatment effect for anyone who fished one of the following permits before 1989: Kodiak setnet, Kodiak purse seine, Cook Inlet drift net  
3.  Calculate the pounds landed per species and earnings per species for the dominant ~ 15 species from these groups (mostly salmon, shellfish, etc).  Also calculate total annual revenue, mean days fished (landed days) and diversity metrics (species, gear, etc).  
4.  Fit a GLM (logistic regression) to the pre-spill data to identify the best predictors of classifying exposure.    
5.  With the same model in (4) use propensity score matching to find best matches for individual fisherman in the "pre" & exposed group, to create a control group  
6. Calculate BACI style metrics, similar to the IFQ comparison to look at changes in groups (control, treatment) pre/post 1989

### Take home: there doesn't seem to be a negative effect of revenue on the treatement / exposed group. But pounds (and revenue) of pink (#440) and chum (#450) salmon decreased more so for this group, and species diversity also appears to be affected. People in the not exposed group fished more coho post-EVOS.

```{r, warning = FALSE, message=FALSE, results = "hide", echo=FALSE}

# Don't include 1989, but include before / after
dat = cfec[cfec$year %in% c(1985:1988, 1990:1991),]
dat$day = substr(dat$landdate, 6, 10)

# Create an indicator based on the salmon / shellfish araes that were closed in PWS
# If someone was fishing in the area before EVOS, they are affected
#dat$exposed = 0
#dat$exposed[dat$stat6%in%c(22710, 22720, 22730, 22610, 22620, 22630, 22640, 
#  22650, 22510, 22520, 22530) & dat$year < 1989] = 1

# For the new analysis (June 2016), based on Eric's plots of the salmon fisherment, the 
# most obvious 'exposed' group are people who had the following permits:
# Kodiak setnet, Kodiak purse seine, Cook Inlet drift net. PWS set net could potentially be incl,
# but participation in 1989 dropped from 560 to 500 people -- didn't go to 0
dat$exposed = 0
dat$exposed[which(dat$p_fshy%in%c("S 03H","S 04K","S 01K") & dat$year <= 1989)] = 1

# Exposed variable is the response, and we want to predict based on traits
dat.ev = group_by(dat, p_holder, year) %>% 
  summarize(tot.revenue.ind = sum(g_earn), 
    tot.rev.230 = sum(g_earn[specn==230]),
    tot.rev.420 = sum(g_earn[specn==420]), 
    tot.rev.450 = sum(g_earn[specn==450]), 
    tot.rev.430 = sum(g_earn[specn==430]), 
    tot.rev.440 = sum(g_earn[specn==440]), 
    tot.rev.410 = sum(g_earn[specn==410]),
    tot.rev.710 = sum(g_earn[specn==710]),
    tot.rev.910 = sum(g_earn[specn==910]), 
    tot.rev.400 = sum(g_earn[specn==400]), 
    tot.rev.168 = sum(g_earn[specn==168]), 
    tot.rev.270 = sum(g_earn[specn==270]),    
    tot.rev.130 = sum(g_earn[specn==130]),
    tot.rev.931 = sum(g_earn[specn==931]),
    tot.rev.830 = sum(g_earn[specn==830]),
    tot.rev.520 = sum(g_earn[specn==520]),    
    tot.pounds.ind = sum(g_pounds), 
    tot.pnds.230 = sum(g_pounds[specn==230]),
    tot.pnds.420 = sum(g_pounds[specn==420]), 
    tot.pnds.450 = sum(g_pounds[specn==450]), 
    tot.pnds.430 = sum(g_pounds[specn==430]), 
    tot.pnds.440 = sum(g_pounds[specn==440]), 
    tot.pnds.410 = sum(g_pounds[specn==410]),
    tot.pnds.710 = sum(g_pounds[specn==710]),
    tot.pnds.910 = sum(g_pounds[specn==910]), 
    tot.pnds.400 = sum(g_pounds[specn==400]), 
    tot.pnds.168 = sum(g_pounds[specn==168]), 
    tot.pnds.270 = sum(g_pounds[specn==270]),    
    tot.pnds.130 = sum(g_pounds[specn==130]),
    tot.pnds.931 = sum(g_pounds[specn==931]),
    tot.pnds.830 = sum(g_pounds[specn==830]),
    tot.pnds.520 = sum(g_pounds[specn==520]),    
    eff.spec.div = simp.div(table(specn)), 
    eff.gear.div = simp.div(table(gearn)),
    numDays = length(unique(day)),
    exposed = max(exposed), 
    ind = ifelse(year[1] < 1989, "Pre", "Post")) %>% 
    as.data.frame()

# Now summarize over years 
  dat.ev = group_by(dat.ev, p_holder, ind) %>% 
  summarize(tot.revenue.ind = mean(tot.revenue.ind), 
    tot.pounds.ind = mean(tot.pounds.ind), 
    eff.spec.div = mean(eff.spec.div), 
    eff.gear.div = mean(eff.gear.div),
    exposed = max(exposed),
    meanDays = mean(numDays),
    tot.rev.230 = mean(tot.rev.230)/tot.revenue.ind,
    tot.rev.420 = mean(tot.rev.420)/tot.revenue.ind, 
    tot.rev.450 = mean(tot.rev.450)/tot.revenue.ind, 
    tot.rev.430 = mean(tot.rev.430)/tot.revenue.ind, 
    tot.rev.440 = mean(tot.rev.440)/tot.revenue.ind, 
    tot.rev.410 = mean(tot.rev.410)/tot.revenue.ind,
    tot.rev.710 = mean(tot.rev.710)/tot.revenue.ind,
    tot.rev.910 = mean(tot.rev.910)/tot.revenue.ind, 
    tot.rev.400 = mean(tot.rev.400)/tot.revenue.ind, 
    tot.rev.168 = mean(tot.rev.168)/tot.revenue.ind, 
    tot.rev.270 = mean(tot.rev.270)/tot.revenue.ind,    
    tot.rev.130 = mean(tot.rev.130)/tot.revenue.ind,
    tot.rev.931 = mean(tot.rev.931)/tot.revenue.ind,
    tot.rev.830 = mean(tot.rev.830)/tot.revenue.ind,
    tot.rev.520 = mean(tot.rev.520)/tot.revenue.ind,    
    tot.pnds.230 = mean(tot.pnds.230)/tot.pounds.ind,
    tot.pnds.420 = mean(tot.pnds.420)/tot.pounds.ind, 
    tot.pnds.450 = mean(tot.pnds.450)/tot.pounds.ind, 
    tot.pnds.430 = mean(tot.pnds.430)/tot.pounds.ind, 
    tot.pnds.440 = mean(tot.pnds.440)/tot.pounds.ind, 
    tot.pnds.410 = mean(tot.pnds.410)/tot.pounds.ind,
    tot.pnds.710 = mean(tot.pnds.710)/tot.pounds.ind,
    tot.pnds.910 = mean(tot.pnds.910)/tot.pounds.ind, 
    tot.pnds.400 = mean(tot.pnds.400)/tot.pounds.ind, 
    tot.pnds.168 = mean(tot.pnds.168)/tot.pounds.ind, 
    tot.pnds.270 = mean(tot.pnds.270)/tot.pounds.ind,    
    tot.pnds.130 = mean(tot.pnds.130)/tot.pounds.ind,
    tot.pnds.931 = mean(tot.pnds.931)/tot.pounds.ind,
    tot.pnds.830 = mean(tot.pnds.830)/tot.pounds.ind,
    tot.pnds.520 = mean(tot.pnds.520)/tot.pounds.ind
    ) %>% 
    as.data.frame()

@@

# We've calculated pounds and revenue of each of the top 15 spp -- including
# both is potentially problematic though because they're correlated ,as expected.
# So for now we'll just focus on revenue. 
dat.ev.rf = dat.ev.rf[,-c(4, 24:38, 40)]

@@

dat.ev.rf[,c(3:5, 23)] = log(dat.ev.rf[,c(3:5, 23)] + 1)
# Select variables that can help explain variation in exposure 
m.glm = glm(exposed ~ ., data = dat.ev.rf[dat.ev.rf$ind=="Pre", -c(1,2)], family = "binomial")
stepAIC(m.glm, direction="both")

bestModel = glm(exposed ~ tot.revenue.ind + eff.spec.div + eff.gear.div + meanDays + 
    tot.rev.420 + tot.rev.450 + tot.rev.430 + tot.rev.410 + 
    tot.rev.710 + tot.rev.400 + tot.rev.168 + tot.rev.270, data = 
    dat.ev.rf[dat.ev.rf$ind == "Pre", ])

m.out <- matchit(exposed ~ tot.revenue.ind + eff.spec.div + eff.gear.div + meanDays + 
    tot.rev.420 + tot.rev.450 + tot.rev.430 + tot.rev.410 + 
    tot.rev.710 + tot.rev.400 + tot.rev.168 + tot.rev.270, method = "nearest", data = dat.ev.rf[dat.ev.rf$ind == "Pre", ])

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
